---
title: "Solutions to Problem 1 & 2"
author: "Yizhi Ma"
date: 2018-09-21
output: github_document
---

```{r, include = FALSE}
library(tidyverse)
library(knitr)
```
 
 
# Problem 1

```{r}
## Set up the data frame for problem 1

  unif_sample = runif(10, min =0, max = 5)

  logi_sample = unif_sample > 2
  
  chara_sample = as.character(logi_sample)
  
  factor_sample = factor(chara_sample)

  prob1_df = data.frame(unif_sample, logi_sample, chara_sample, factor_sample)
  
kable(prob1_df, caption = "Data Frame for Problem 1")
```

The mean works for the first two columns of the data frame, and dosen't work for the character vector and factor vector. This is because we can only calculate mean for numeric or logical variable.  

```{r}
## Convert variable to numeric
logi_sample_to_num = as.numeric(logi_sample)  
chara_sample_to_num = as.numeric(chara_sample)  
factor_sample_to_num = as.numeric(factor_sample)  
```

After the conversion, the TRUE/FALSE in logical variable is converted to 1/0, and the TRUE/FALSE in factor variable is converted to 2/1, character variable is not applicable for this conversion.

```{r}
## Two-step conversion
character_to_factor_to_numeric = as.numeric(as.factor(chara_sample))
factor_to_character_to_numeric = as.numeric(as.character(factor_sample))
```
Obviously, the character_to_factor_to_numeric conversion is successful, the TRUE/FALSE are converted to 2/1. However, the factor_to_character_to_numeric conversion failed because character cannot be converted to numeric variable.

# Problem 2

```{r}
## Set up the data fraame for problem 2
prob2_df = tibble(
  x = rnorm(1000, mean = 0, sd = 1),
  y = rnorm(1000, mean = 0, sd = 1),    
  vec_logi = x+y > 0,
  vec_num = as.numeric(vec_logi),
  vec_factor = as.factor(vec_logi)
)
```

The size of this dataset is 1000 rows by 2 columns.  
The mean and median of x are both 0.  
50.3% of cases for which the logical vector is TRUE.

```{r}
## Scatter plot of x and y, color using logical vector
ggplot(prob2_df, aes(x = x, y = y, color = vec_logi)) + geom_point()

## Scatter plot of x and y, color using numerical vector
ggplot(prob2_df, aes(x = x, y = y, color = vec_num)) + geom_point()

## Scatter plot of x and y, color using factor vector
ggplot(prob2_df, aes(x = x, y = y, color = vec_factor)) + geom_point()
```